Introducing Distributed Dynamic Data-intensive (D3) Science: Understanding Applications and Infrastructure

### 2016-02-14 3dpas_intro.tex
introduction - reduced and refined.

The landscape of scientific and enterprise computing is being fundamentally
altered by the prodigious volumes of data.  % Not only are the volumes of data
% produced by scientific applications increasing rapidly, there is also a
% corresponding increase in the amount of data that needs to be processed and
% analyzed.
With the increase in data volumes, types, and sources of data, a concomitant
development has been an increase in the sophistication and complexity of the
infrastructure needed to support these requirements. For example, data are often
naturally distributed and have to be processed in situ; occasionally data that
are localized have to be distributed in order to be processed, either for
reasons of performance or for collaborative analytics.  The scale and
significance of varying data source/destination and transfer/production rates
makes temporal variations important, which in turn eventually impact the design
and development of infrastructure to support them: resource scheduling, data
placement, and transfer decisions that were statically determinable at small
scales, emerge as dynamic and important considerations at large scales of data
and distribution.

@@

The challenges of distributed data-intensive applications span beyond the
management and curation of large volumes of data. A critical requirement is to
transform the data into information en route to assembling knowledge. Among
other things, the transformation from data to information requires extensive
amounts of computational processing of the data; thus the data-intensive nature
of the data-intensive science problem is strongly coupled with compute-intensive
requirements.

@@

Analyzing the applications along an analytical space spanned by the three ``D''s
of data-intensive, distributed, and dynamic provides a general and effective
framework.  Subjecting a range of otherwise distinct applications to a common
analytical framework suggests to infrastructure scientists and
cyberinfrastructure experts where effort is needed
and is likely to yield maximum returns.  The diverse set of requirements to support
D3 scenarios points to the need for a careful understanding of D3
characteristics as well as sophisticated, extensible, and skillfully architected
infrastructure.

As D3 applications evolve, our analysis can provide guidance on where to focus
effort on infrastructure.  Last but not least, the advantages of providing a
common but extensible vocabulary that can be used by the entire community,
application end-users, developers and infrastructure scientists alike, in an
emerging field cannot be overstated. We believe that such vocabulary and
terminology has been missing from the practice of distributed
cyberinfrastructure in general, with negative consequences.

>>>>

% Not only are the volumes of data
% produced by scientific applications increasing rapidly, there is also a
% corresponding increase in the amount of data that needs to be processed and
% analyzed.

The landscape of scientific and enterprise computing is being fundamentally
altered by the prodigious volumes of data.  Occasionally data are often
naturally distributed and have to be processed near source; sometimes data that
are localized have to be distributed in order to be processed, either for
reasons of performance or for collaborative analytics. Further, the scale of
varying data source and destination, as well as the transfer of production rates
makes temporal variations in data properties increasingly important.  With an
increase in the importance of distributed and temporal properties of data,
systems and infrastructure aspects such as resource scheduling, data placement,
and transfer decisions that were statically determinable at small scales, emerge
as dynamic and important considerations at large scales of data and
distribution.

@@

% The challenges of distributed data-intensive applications span beyond the
% management and curation of large volumes of data. A critical requirement is to
% transform the data into information en route to assembling knowledge. Among
% other things, the transformation from data to information requires extensive
% amounts of computational processing of the data; thus the data-intensive nature
% of the data-intensive science problem is strongly coupled with compute-intensive
% requirements.

@@

The analytical space spanned by the three ``D''s of data-intensive, distributed,
and dynamic provides a common vocabulary and effective framework to analyze
applications and infrastructure alike.  Subjecting a range of otherwise distinct
applications to a common analytical framework provides an opportunity to search
for similarities across applications and understand core differences.  The
diverse % set of requirements to support
D3 scenarios points to the need for a careful understanding of D3
characteristics as well as sophisticated, extensible, and skillfully architected
infrastructure.  A common framework thus informs infrastructure scientists and
cyberinfrastructure experts where effort is likely to yield maximum returns.
The advantages of providing a common but extensible vocabulary that can be used
by the entire community, application end-users, developers and infrastructure
scientists alike, in an emerging field cannot be overstated. We believe that
such vocabulary and terminology has been missing from the practice of
distributed cyberinfrastructure in general, with negative consequences.

% As D3 applications
% evolve, our analysis can provide guidance on where to focus effort on
% infrastructure.

% Analyzing the applications along an analytical space spanned by the three ``D''s
% of data-intensive, distributed, and dynamic provides a general and effective
% framework.  Subjecting a range of otherwise distinct applications to a common
% analytical framework suggests to infrastructure scientists and
% cyberinfrastructure experts where effort is needed and is likely to yield
% maximum returns.  The diverse set of requirements to support D3 scenarios points
% to the need for a careful understanding of D3 characteristics as well as
% sophisticated, extensible, and skillfully architected infrastructure.
